### 2022-01-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_transfer@8ddafbb3de0b99e96ee7fb253b6d3505e2c1f382 🚀
Deploying to gh-pages from @ PoisotLab/ms_metaweb_transfer@296421f2fd91597c6d2c5332a653469ae1a20b64 🚀

There are two core challenges we are faced with in furthering our
understanding of ecological networks across space, particularly at
macro-ecologically relevant scales (\emph{e.g.} Trøjelsgaard \& Olesen,
2016). First, ecological networks within a location are difficult to
sample properly (Jordano, 2016a, 2016b), resulting in a widespread
``Eltonian shortfall'' (Hortal et al., 2015), \emph{i.e.} a lack of
knowledge about inter- and intra- specific relationships. This first
challenge has been, in large part, addressed by the recent emergence of
a suite of methods aiming to predict interactions within \emph{existing}
networks, many of which are reviewed in Strydom, Catchen, et al. (2021).
Second, recent analyses based on collected data (Poisot, Bergeron, et
al., 2021) or metadata (Cameron et al., 2019) highlight that ecological
networks are currently studied in a biased subset of space and
bioclimates, which impedes our ability to generalize any local
understanding of network structure. Meaning that, although the framework
to address incompleteness \emph{within} networks exists, there would
still be regions for which, due to a \emph{lack} of local interaction
data, we are unable to infer potential species interactions. Having a
general solution for inferring \emph{potential} interactions (despite
the unavailability of interaction data) could be the catalyst for
significant breakthroughs in our ability to start thinking about species
interaction networks over large spatial scales. In a recent overview of
the field of ecological network prediction, Strydom, Catchen, et al.
(2021) identified two challenges of interest to the prediction of
interactions at large scales. First, there is a relative scarcity of
relevant data in most places globally -- paradoxically, this restricts
our ability to infer interactions to locations where inference is
perhaps the least required; second, accurate predictions often demand
accurate predictors, and the lack of methods that can leverage small
amount of data is a serious impediment to our predictive ability
globally.

@@

In fig.~\ref{fig:concept}, where we provide a methodological overview
based on learning the embedding of a metaweb of trophic interactions for
European mammals (known interactions; Maiorano et al., 2020a, 2020b)
and, based on phylogenetic relationships between mammals globally
(\emph{i.e.}, phylogenetic tree Upham et al., 2019), infer a metaweb for
the Canadian mammalian species pool (using only a species list
\emph{i.e.} interactions are `unknown' in this instance). Following the
definition of Dunne (2006), a metaweb is a network analogue to the
regional species pool; specifically, it is an inventory of all
\emph{potential} interactions between species from a spatially delimited
area (and so captures the \(\gamma\) diversity of interactions). The
metaweb is, therefore, \emph{not} a prediction of the food web at a
specific locale within the spatial area it covers, and will have a
different structure (notably by having a larger connectance; see
\emph{e.g.} Wood et al., 2015). These local food webs (which captures
the \(\alpha\) diversity of interactions) are a subset of the metaweb's
species and interactions, and have been called ``metaweb realizations''
(Poisot et al., 2015). Differences between local food web and their
metaweb are due to chance, species abundance and co-occurrence, local
environmental conditions, and local distribution of functional traits,
among others.

Because the metaweb represents the joint effect of functional,
phylogenetic, and macroecological processes (Morales-Castilla et al.,
2015), it holds valuable ecological information. Specifically, it is the
``upper bounds'' on what the composition of the local networks can be
(see \emph{e.g.} McLeod et al., 2021). These local networks, in turn,
can be reconstructed given appropriate knowledge of local species
composition, providing information on structure of food webs at finer
spatial scales. This has been done for example for
tree-galler-parasitoid systems (Gravel et al., 2018), fish trophic
interactions (Albouy et al., 2019), tetrapod trophic interactions
(O'Connor et al., 2020), and crop-pest networks (Grünig et al., 2020).
Whereas the original metaweb definition, and indeed most past uses of
metawebs, was based on the presence/absence of interactions, we focus on
\emph{probabilistic} metawebs where interactions are represented as the
chance of success of a Bernoulli trial (see \emph{e.g.} Poisot et al.,
2016); therefore, not only does our method recommend interactions that
may exist, it gives each interaction a score, allowing us to properly
weigh them.

Our case study shows that phylogenetic transfer learning is an effective
approach to the generation of probabilistic metawebs. This showcases
that although the components (species) that make up the Canadian and
European communities may be \emph{minimally} shared (the overall species
overlap is less than 4\%), if the medium (proxy space) selected in the
transfer step is biologically plausible, we can still effectively learn
from the known network and make biologically relevant predictions of
interactions. Indeed, as we detail in the results, when validated
against known but fractional data of trophic interactions between
Canadian mammals, our model achieves a predictive accuracy of
approximately 91\%. It should be reiterated that the framework presented
in fig.~\ref{fig:concept} is amenable to changes \emph{e.g.} it is
possible to use distinct trees if working with distinct clades (such as
pollination networks) or, alternatively, the measure of similarity may
not be phylogeny, and can be replaced by information on foraging
(Beckerman et al., 2006), cell-level mechanisms (Boeckaerts et al.,
2021), or a combination of traits and phylogenetic structure (Stock,
2021). Most importantly, although we focus on a trophic system, it is an
established fact that different (non-trophic) interactions do themselves
interact with and influence the outcome of trophic interactions (see
\emph{e.g.} Kawatsu et al., 2021; Kéfi et al., 2012). Future development
of metaweb inference techniques should cover the prediction of multiple
interaction types.

\hypertarget{data-used-for-the-case-study}{%
\section{Data used for the case
study}\label{data-used-for-the-case-study}}

We use data from the European metaweb assembled by Maiorano et al.
(2020a). This was assembled using data extracted from scientific
literature (including published papers, books, and grey literature) from
the last 50 years and includes all terrestrial tetrapods (mammals,
breeding birds, reptiles and amphibians) occurring on the European
sub-continent (and Turkey) - with the caveat that only species
introduced in historical times and currently naturalized being included.
The European metaweb was filtered using the Global Biodiversity
Information Facility (GBIF) taxonomic backbone (GBIF Secretariat, 2021)
so as to contain only terrestrial and semi-aquatic mammals. As all
species had valid matches to the GBIF taxonomy it was used as the
backbone for the remaining reconciliation steps namely, the mammalian
consensus supertree by Upham et al. (2019) (which is used for the
knowledge transfer step) and for the Canadian species list---which was
extracted from the International Union for Conservation of Nature (IUCN)
checklist, and corresponds to the same selection criteria that was
applied by Maiorano et al. (2020a) in the European metaweb.

After taxonomic cleaning and reconciliation the mammalian European
metaweb has 260 species, and the Canadian species pool 163; of these, 17
(about 4\% of the total) are shared, and 89 species from Canada (54\%)
had at least one congeneric species in Europe. The similarity for both
species pools predictably increases with higher taxonomic order, with
19\% of shared genera, 47\% of shared families, and 75\% of shared
orders; for the last point, Canada and Europe each had a single unique
order (\emph{Didelphimorphia} for Canada, \emph{Erinaceomorpha} for
Europe).

@@

The core point of our method is the transfer of knowledge of a known
ecological network, in order to predict interactions between species
from another location for which the network is unknown (or partially
known). In fig.~\ref{fig:concept} the grey text boxes give a high-level
overview of the approach; in the example around which this manuscript is
built The method we develop is, ecologically speaking, a ``black box,''
\emph{i.e.} an algorithm that can be understood mathematically, but
whose component parts are not always directly tied to ecological
processes. There is a growing realization in machine learning that
(unintentional) black box algorithms are not necessarily a bad thing
(Holm, 2019), as long as their constituent parts can be examined (which
is the case with our method). But more importantly, data hold more
information than we might think; as such, even algorithms that are
disconnected from a model can make correct guesses most of the time
(Halevy et al., 2009); in fact, in an instance of ecological forecasting
of spatio-temporal systems, model-free approaches (\emph{i.e.} drawing
all of their information from the data) outperformed model-informed ones
(Perretti et al., 2013).

@@

The first step in transfer learning is to learn the structure of the
original dataset. In order to do so, we rely on an approach inspired
from representational learning, where we learn a \emph{representation}
of the metaweb (in the form of the latent subspaces), rather than a list
of interactions (species \emph{a} eats \emph{b}). This approach is
conceptually different from other metaweb-scale predictions (\emph{e.g.}
Albouy et al., 2019), in that the metaweb representation is easily
transferable. Specifically, we use Random Dot Product Graph model
(hereafter RDPG; S. J. Young \& Scheinerman, 2007) to create a number of
latent variables that can be combined into an approximation of the
network adjacency matrix. RDPG is known to capture the evolutionary
backbone of food webs (Dalla Riva \& Stouffer, 2016), resulting in
strong phylogenetic signal in RDPG results; in other words, the latent
variables of an RDPG can be mapped onto a phylogenetic tree, and
phylogenetically similar predators should share phylogenetically similar
preys. In addition, recent advances show that the latent variables
produced this way can be used to predict \emph{de novo} interactions.
Interestingly, the latent variables do not need to be produced by
decomposing the network itself; in a recent contribution, Runghen et al.
(2021) showed that deep artificial neural networks are able to
reconstruct the left and right subspaces of an RDPG, in order to predict
human movement networks from individual/location metadata and opens up
the possibility of using additional metadata as predictors.

@@

In order to transfer the knowledge from the European metaweb to the
Canadian species pool, we performed ancestral character estimation using
a Brownian motion model, which is a conservative approach in the absence
of strong hypotheses about the nature of phylogenetic signal in the
network decomposition (Litsios \& Salamin, 2012). This uses the
estimated feature vectors for the European mammals to create a state
reconstruction for all species (conceptually something akin to a
trait-based mammalian phylogeny using latent generality and
vulnerability traits) and allows us to impute the missing (latent) trait
data for the Canadian species that are not already in the European
network; as we are focused on predicting contemporary interactions, we
only retained the values for the tips of the tree. We assumed that all
traits (\emph{i.e.} the feature vectors for the left and right
subspaces) were independent, which is a reasonable assumption as every
trait/dimension added to the t-SVD has an \emph{additive} effect to the
one before it. Note that the Upham et al. (2019) tree itself has some
uncertainty associated to inner nodes of the phylogeny. In this case
study, we have decided to not propagate this uncertainty, as it would
complexify the process. The Brownian motion algorithm returns the
\emph{average} value of the trait, and its upper and lower bounds.
Because we do not estimate other parameters of the traits'
distributions, we considered that every species trait is represented as
a uniform distribution between these bounds. The choice of the uniform
distribution was made because the algorithm returns a minimum and
maximum point estimate for the value, and given this information, the
uniform distribution is the one with maximum entropy. Had all mean
parameters estimates been positive, the exponential distribution would
have been an alternative, but this is not the case for the subspaces of
an RDPG. In order to examine the consequences of the choice of
distribution, we estimated the variance per latent variable per node to
use a Normal distribution; as we show in Supp. Mat. 2, this decision
results in dramatically over-estimating the number and probability of
interactions, and therefore we keep the discussions in the main text to
the uniform case. The inferred left and right subspaces for the Canadian
species pool (\(\hat{\mathscr{L}}\) and \(\hat{\mathscr{R}}\)) have
entries that are distributions, representing the range of values for a
given species at a given dimension.

These objects represent the transferred knowledge, which we can use for
prediction of the Canadian metaweb.

@@

First, we extracted the subgraph corresponding to the 17 species shared
between the European and Canadian pools and replaced these interactions
with a probability of 0 (non-interaction) or 1 (interaction), according
to their value in the European metaweb. This represents a minute
modification of the inferred network (about 0.8\% of all species pairs
from the Canadian web), but ensures that we are directly re-using
knowledge from Europe.

@@

Because the confidence intervals on the inferred trait space are
probably over-estimates, we decided to apply a thresholding step to the
interactions after the data inflation (see fig.~\ref{fig:thresholds}
showing the effect of varying the cutoff on \(P(i \rightarrow j)\)).
Cirtwill \& Hambäck (2021) proposed a number of strategies to threshold
probabilistic networks. Their methods assume the underlying data to be
tag-based sequencing, which represents interactions as co-occurrences of
predator and prey within the same tags; this is conceptually identical
to our Bernoulli-trial based reconstruction of a probabilistic network.
We performed a full analysis of the effect of various cutoffs, and as
they either resulted in removing too few interactions, or removing
enough interactions that species started to be disconnected from the
network, we set this threshold for a probability equivalent to 0 to the
largest possible value that still allowed all species to have at least
one interaction with a non-zero probability. The need for this slight
deviation from the Cirtwill \& Hambäck (2021) method highlights the need
for additional development on network thresholding.

@@

The t-SVD embedding is able to learn relevant ecological features for
the network. fig.~\ref{fig:degree} shows that the first rank correlates
linearly with generality and vulnerability (Schoener, 1989), \emph{i.e.}
the number of preys and predators for each species. Importantly, this
implies that a rank 1 approximation represents the configuration model
for the metaweb, \emph{i.e.} a set of random networks generated from a
given degree sequence (Park \& Newman, 2004). Accounting for the
probabilistic nature of the degrees, the rank 1 approximation also
represents the \emph{soft} configuration model (van der Hoorn et al.,
2018). Both models are maximum entropy graph models (Garlaschelli et
al., 2018), with sharp (all network realizations satisfy the specified
degree sequence) and soft (network realizations satisfy the degree
sequence on average) local constraints, respectively. The (soft)
configuration model is an unbiased random graph model widely used by
ecologists in the context of null hypothesis significance testing of
network structure (\emph{e.g.} Bascompte et al., 2003) and can provide
informative priors for Bayesian inference of network structure
(\emph{e.g.} J.-G. Young et al., 2021). It is noteworthy that for this
metaweb, the relevant information was extracted at the first rank.
Because the first rank corresponds to the leading singular value of the
system, the results of fig.~\ref{fig:degree} have a straightforward
interpretation: degree-based processes are the most important in
structuring the mammalian food web.

@@

One important aspect in which Europe and Canada differ (despite their
comparable bioclimatic conditions) is the degree of the legacy of human
impacts, which have been much longer in Europe. Nenzén et al. (2014)
showed that even at small scales (the Iberian peninsula), mammal food
webs retain the signal of both past climate change and human activity,
even when this human activity was orders of magnitude less important
than it is now. Similarly, Yeakel et al. (2014) showed that changes in
human occupation over several centuries can lead to food web collapse.
Megafauna in particular seems to be very sensitive to human arrival
(Pires et al., 2015). In short, there is well-substantiated support for
the idea that human footprint affects more than the risk of species
extinction (Marco et al., 2018), and can lead to changes in interaction
structure. Yet, owing to the inherent plasticity of interactions, there
have been documented instances of food webs undergoing rapid
collapse/recovery cycles over short periods of time (Pedersen et al.,
2017). The embedding of a network, in a sense, embeds its
macro-evolutionary history, especially as RDPG captures ecological
signal (Dalla Riva \& Stouffer, 2016); at this point, it is important to
recall that a metaweb is intended as a catalogue of all potential
interactions, which should then be filtered (Morales-Castilla et al.,
2015). In practice (and in this instance) the reconstructed metaweb will
predict interactions that are plausible based on the species'
evolutionary history, however some interactions would/would not be
realized due to human impact.

Dallas et al. (2017) suggested that most links in ecological networks
may be cryptic, \emph{i.e.} uncommon or otherwise hard to observe. This
argument essentially echoes Jordano (2016b): the sampling of ecological
interactions is difficult because it requires first the joint
observation of two species, and then the observation of their
interaction. In addition, it is generally expected that weak or rare
links would be more common in networks (Csermely, 2004), compared to
strong, persistent links; this is notably the case in food chains,
wherein many weaker links are key to the stability of a system (Neutel
et al., 2002). In the light of these observations, the results in
fig.~\ref{fig:inflation} are not particularly surprising: we expect to
see a surge in these low-probability interactions under a model that has
a good predictive accuracy. Because the predictions we generate are by
design probabilistic, then one can weigh these rare links appropriately.
In a sense, that most ecological interactions are elusive can call for a
slightly different approach to sampling: once the common interactions
are documented, the effort required in documenting each rare interaction
may increase exponentially. Recent proposals suggest that machine
learning algorithms, in these situations, can act as data generators
(Hoffmann et al., 2019): in this perspective, high quality observational
data can be supplemented with synthetic data coming from predictive
models, which increases the volume of information available for
inference. Indeed, Strydom, Catchen, et al. (2021) suggested that
knowing the metaweb may render the prediction of local networks easier,
because it fixes an ``upper bound'' on which interactions can exist;
indeed, with a probabilistic metaweb, we can consider that the metaweb
represents an aggregation of informative priors on the interactions.

Related to the last point, Cirtwill et al. (2019) showed that network
inference techniques based on Bayesian approaches would perform far
better in the presence of an interaction-level informative prior; the
desirable properties of such a prior would be that it is expressed as a
probability, preferably representing a Bernoulli event, the value of
which would be representative of relevant biological processes
(probability of predation in this case). We argue that the probability
returned at the very last step of our framework may serve as this
informative prior; indeed, the output of our analysis can be used in
subsequent steps, also possibly involving expert elicitation to validate
some of the most strongly recommended interactions. One important
\emph{caveat} to keep in mind when working with interaction inference is
that interactions can never really be true negatives (in the current
state of our methodological framework and data collection limitations);
this renders the task of validating a model through the usual
application of binary classification statistics very difficult (although
see Strydom, Catchen, et al., 2021 for a discussion of alternative
suggestions). The other way through which our framework can be improved
is by substituting the predictors that are used for transfer. For
example, in the presence of information on species traits that are known
to be predictive of species interactions, one might want to rely on
functional rather than phylogenetic distances -- in food webs, body size
(and allometrically related variables) has been established as such a
variable (Brose et al., 2006); the identification of relevant functional
traits is facilitated by recent methodological developments (Rosado et
al., 2013). It should be noted that Xing \& Fayle (2021) highlight
phylogenetic relatedness as one of the core components of network
comparison at the global scale. In this case study, we have embedded the
original metaweb using t-SVD, because it lends itself to an RDPG
reconstruction, which is known to capture the consequences of
evolutionary processes (Dalla Riva \& Stouffer, 2016); this being said,
there are other ways to embed graphs (Arsov \& Mirceva, 2019; Cai et
al., 2017; Cao et al., 2019), which can be used as alternatives.

As Herbert (1965) rightfully pointed out, ``{[}y{]}ou can't draw neat
lines around planet-wide problems''; in this regard, our approach (and
indeed, any inference of a metaweb at large scales) must contend with
several interesting and interwoven families of problems. The first is
the limit of the metaweb to embed and transfer. If the initial metaweb
is too narrow in scope, notably from a taxonomic point of view, the
chances of finding another area with enough related species to make a
reliable inference decreases; this would likely be indicated by large
confidence intervals during ancestral character estimation, but the lack
of well documented metawebs is currently preventing the development of
more concrete guidelines. The question of phylogenetic relatedness and
dispersal is notably true if the metaweb is assembled in an area with
mostly endemic species, and as with every predictive algorithm, there is
room for the application of our best ecological judgement. Conversely,
the metaweb should be reliably filled, which assumes that the \(S^2\)
interactions in a pool of \(S\) species have been examined, either
through literature surveys or expert elicitation. Supp. Mat. 1 provides
some guidance as to the type of sampling effort that should be
prioritized. While RDPG was able to maintain very high predictive power
when interactions were missing, the addition of false positive
interactions was immediately detected; this suggests that it may be
appropriate to err on the side of ``too many'' interactions when
constructing the initial metaweb to be transferred. The second series of
problems are related to determining which area should be used to infer
the new metaweb in, as this determines the species pool that must be
used. In our application, we focused on the mammals of Canada. The
upside of this approach is that information at the country level is
likely to be required by policy makers and stakeholders for their
biodiversity assessment, as each country tends to set goals at the
national level (Buxton et al., 2021) for which quantitative instruments
are designed (Turak et al., 2017), with specific strategies often
enacted at smaller scales (Ray et al., 2021). And yet, we do not really
have a satisfying answer to the question of ``where does a food web
stop?''; the current most satisfying solutions involve examining the
spatial consistency of network area relationships (Fortin et al., 2021;
see \emph{e.g.} Galiana et al., 2018, 2019, 2021), which is of course
impossible in the absence of enough information about the network
itself. This suggests that an \emph{a posteriori} refinement of the
results may be required, based on a downscaling of the metaweb. The
final family of problems relates less to the availability of data or
quantitative tools, and more to the praxis of spatial ecology. Operating
under the context of national divisions, in large parts of the world,
reflects nothing more than the legacy of settler colonialism. Indeed,
the use of ecological data is not an apolitical act (Nost \& Goldstein,
2021), as data infrastructures tend to be designed to answer questions
within national boundaries, and their use both draws upon and reinforces
territorial statecraft; as per Machen \& Nost (2021), this is
particularly true when the output of ``algorithmic thinking''
(\emph{e.g.} relying on machine learning to generate knowledge) can be
re-used for governance (\emph{e.g.} enacting conservation decisions at
the national scale). We therefore recognize that methods such as we
propose operate under the framework that contributed to the ongoing
biodiversity crisis (Adam, 2014), reinforced environmental injustice
(Choudry, 2013; Domínguez \& Luoma, 2020), and on Turtle Island
especially, should be replaced by Indigenous principles of land
management (Eichhorn et al., 2019; No'kmaq et al., 2021). As we see
AI/ML being increasingly mobilized to generate knowledge that is lacking
for conservation decisions (\emph{e.g.} Lamba et al., 2019; Mosebo
Fernandes et al., 2020), our discussion of these tools need to go beyond
the technical, and into the governance consequences they can have.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Adam2014EleTre}{}%
Adam, R. (2014). \emph{Elephant treaties: The Colonial legacy of the
biodiversity crisis}. UPNE.

>>>>

There are two core challenges we are faced with in furthering our
understanding of ecological networks across space, particularly at
macro-ecologically relevant scales (\emph{e.g.} Trøjelsgaard \& Olesen,
2016). First, ecological networks within a location are difficult to
sample properly (Jordano, 2016a, 2016b), resulting in a widespread
``Eltonian shortfall'' (Hortal et al., 2015), \emph{i.e.} a lack of
knowledge about inter- and intra- specific relationships. This first
challenge has been, in large part, addressed by the recent emergence of
a suite of methods aiming to predict interactions within \emph{existing}
networks, many of which are reviewed in Strydom, Catchen, et al. (2021).
Second, recent analyses based on collected data (Poisot, Bergeron, et
al., 2021) or metadata (Cameron et al., 2019) highlight that ecological
networks are currently studied in a biased subset of space and
bioclimates, which impedes our ability to generalize any local
understanding of network structure. Meaning that, although the framework
to address incompleteness \emph{within} networks exists, there would
still be regions for which, due to a \emph{lack} of local interaction
data, we are unable to infer potential species interactions.

@@

In fig.~\ref{fig:concept}, we provide a methodological overview based on
learning the embedding of a metaweb of trophic interactions for European
mammals (known interactions; Maiorano et al., 2020a, 2020b) and, based
on phylogenetic relationships between mammals globally (\emph{i.e.},
phylogenetic tree Upham et al., 2019), infer a metaweb for the Canadian
mammalian species pool (using only a species list \emph{i.e.}
interactions are treated as unknown in this instance). Our case study
shows that phylogenetic transfer learning is an effective approach to
the generation of probabilistic metawebs. This showcases that although
the components (species) that make up the Canadian and European
communities may be \emph{minimally} shared (the overall species overlap
is less than 4\%), if the medium (proxy space) selected in the transfer
step is biologically plausible, we can still effectively learn from the
known network and make biologically relevant predictions of
interactions. Indeed, as we detail in the results, when validated
against known but fractional data of trophic interactions between
Canadian mammals, our model achieves a predictive accuracy of
approximately 91\%.

@@

The core point of our method is the transfer of knowledge of a known
ecological network to predict interactions between species for another
location for which the network is unknown (or partially known) and is
summarized in the grey text boxes in fig.~\ref{fig:concept}. The method
we develop is, ecologically speaking, a ``black box,'' \emph{i.e.} an
algorithm that can be understood mathematically, but whose component
parts are not always directly tied to ecological processes. There is a
growing realization in machine learning that (unintentional) black box
algorithms are not necessarily a bad thing (Holm, 2019), as long as
their constituent parts can be examined (which is the case with our
method). But more importantly, data hold more information than we might
think; as such, even algorithms that are disconnected from a model can
make correct guesses most of the time (Halevy et al., 2009); in fact, in
an instance of ecological forecasting of spatio-temporal systems,
model-free approaches (\emph{i.e.} drawing all of their information from
the data) outperformed model-informed ones (Perretti et al., 2013).

\hypertarget{data-used-for-the-case-study}{%
\subsection{Data used for the case
study}\label{data-used-for-the-case-study}}

We use data from the European metaweb assembled by Maiorano et al.
(2020a). This was assembled using data extracted from scientific
literature (including published papers, books, and grey literature) from
the last 50 years and includes all terrestrial tetrapods (mammals,
breeding birds, reptiles and amphibians) occurring on the European
sub-continent (and Turkey) - with the caveat that only species
introduced in historical times and currently naturalized being included.
The European metaweb was filtered using the Global Biodiversity
Information Facility (GBIF) taxonomic backbone (GBIF Secretariat, 2021)
so as to contain only terrestrial and semi-aquatic mammals. As all
species had valid matches to the GBIF taxonomy it was used as the
backbone for the remaining reconciliation steps namely, the mammalian
consensus supertree by Upham et al. (2019) (which is used for the
knowledge transfer step) and for the Canadian species list---which was
extracted from the International Union for Conservation of Nature (IUCN)
checklist, and corresponds to the same selection criteria that was
applied by Maiorano et al. (2020a) in the European metaweb. After
taxonomic cleaning and reconciliation the European metaweb has 260
species, and the Canadian species pool 163; of these, 17 (about 4\% of
the total) are shared, and 89 species from Canada (54\%) had at least
one congeneric species in Europe. The similarity for both species pools
predictably increases with higher taxonomic order, with 19\% of shared
genera, 47\% of shared families, and 75\% of shared orders; for the last
point, Canada and Europe each had a single unique order
(\emph{Didelphimorphia} for Canada, \emph{Erinaceomorpha} for Europe).

@@

The first step in transfer learning is to learn the structure of the
original dataset. In order to do so, we rely on an approach inspired
from representational learning, where we learn a \emph{representation}
of the metaweb (in the form of the latent subspaces), rather than a list
of interactions (species \emph{a} eats \emph{b}). This approach is
conceptually different from other metaweb-scale predictions (\emph{e.g.}
Albouy et al., 2019), in that the metaweb representation is easily
transferable. Specifically, we use a Random Dot Product Graph model
(hereafter RDPG; S. J. Young \& Scheinerman, 2007) to create a number of
latent variables that can be combined into an approximation of the
network adjacency matrix. RDPG is known to capture the evolutionary
backbone of food webs (Dalla Riva \& Stouffer, 2016), resulting in
strong phylogenetic signal in RDPG results; in other words, the latent
variables of an RDPG can be mapped onto a phylogenetic tree, and
phylogenetically similar predators should share phylogenetically similar
preys. In addition, recent advances show that the latent variables
produced this way can be used to predict \emph{de novo} interactions.
Interestingly, the latent variables do not need to be produced by
decomposing the network itself; in a recent contribution, Runghen et al.
(2021) showed that deep artificial neural networks are able to
reconstruct the left and right subspaces of an RDPG, in order to predict
human movement networks from individual/location metadata and opens up
the possibility of using additional metadata as predictors.

@@

In order to transfer the knowledge from the European metaweb to the
Canadian species pool, we performed ancestral character estimation using
a Brownian motion model, which is a conservative approach in the absence
of strong hypotheses about the nature of phylogenetic signal in the
network decomposition (Litsios \& Salamin, 2012). This uses the
estimated feature vectors for the European mammals to create a state
reconstruction for all species (conceptually something akin to a
trait-based mammalian phylogeny using latent generality and
vulnerability traits) and allows us to impute the missing (latent) trait
data for the Canadian species that are not already in the European
network; as we are focused on predicting contemporary interactions, we
only retained the values for the tips of the tree. We assumed that all
traits (\emph{i.e.} the feature vectors for the left and right
subspaces) were independent, which is a reasonable assumption as every
trait/dimension added to the t-SVD has an \emph{additive} effect to the
one before it. Note that the Upham et al. (2019) tree itself has some
uncertainty associated to inner nodes of the phylogeny. In this case
study we have decided to not propagate this uncertainty as it would
complexify the process. The Brownian motion algorithm returns the
\emph{average} value of the trait, and its upper and lower bounds.
Because we do not estimate other parameters of the traits'
distributions, we considered that every species trait is represented as
a uniform distribution between these bounds. The choice of the uniform
distribution was made because the algorithm returns a minimum and
maximum point estimate for the value, and given this information, the
uniform distribution is the one with maximum entropy. Had all mean
parameters estimates been positive, the exponential distribution would
have been an alternative, but this is not the case for the subspaces of
an RDPG. In order to examine the consequences of the choice of
distribution, we estimated the variance per latent variable per node to
use a Normal distribution; as we show in Supp. Mat. 2, this decision
results in dramatically over-estimating the number and probability of
interactions, and therefore we keep the discussions in the main text to
the uniform case. The inferred left and right subspaces for the Canadian
species pool (\(\hat{\mathscr{L}}\) and \(\hat{\mathscr{R}}\)) have
entries that are distributions, representing the range of values for a
given species at a given dimension. These objects represent the
transferred knowledge, which we can use for prediction of the Canadian
metaweb.

@@

First, we extracted the network corresponding to the 17 species shared
between the European and Canadian pools and replaced these interactions
with a probability of 0 (non-interaction) or 1 (interaction), according
to their value in the European metaweb. This represents a minute
modification of the inferred network (about 0.8\% of all species pairs
from the Canadian web), but ensures that we are directly re-using
knowledge from Europe.

@@

Because the confidence intervals on the inferred trait space are
probably over-estimates, we decided to apply a thresholding step to the
interactions after data inflation (see fig.~\ref{fig:thresholds} showing
the effect of varying the cutoff on \(P(i \rightarrow j)\)). Cirtwill \&
Hambäck (2021) proposed a number of strategies to threshold
probabilistic networks. Their methods assume the underlying data to be
tag-based sequencing, which represents interactions as co-occurrences of
predator and prey within the same tags; this is conceptually identical
to our Bernoulli-trial based reconstruction of a probabilistic network.
We performed a full analysis of the effect of various cutoffs, and as
they either resulted in removing too few interactions, or removing
enough interactions that species started to be disconnected from the
network, we set this threshold for a probability equivalent to 0 to the
largest possible value that still allowed all species to have at least
one interaction with a non-zero probability. The need for this slight
deviation from the Cirtwill \& Hambäck (2021) method highlights the need
for additional development on network thresholding.

@@

Using a transfer learning framework we were able to construct a
probabilistic metaweb and (as as per Dunne, 2006) it is a list of
potential interactions and does not mean that they will exist wherever
the two species co-occur. The t-SVD embedding is able to learn relevant
ecological features for the network. fig.~\ref{fig:degree} shows that
the first rank correlates linearly with generality and vulnerability
(Schoener, 1989), \emph{i.e.} the number of preys and predators for each
species. Importantly, this implies that a rank 1 approximation
represents the configuration model for the metaweb, \emph{i.e.} a set of
random networks generated from a given degree sequence (Park \& Newman,
2004). Accounting for the probabilistic nature of the degrees, the rank
1 approximation also represents the \emph{soft} configuration model (van
der Hoorn et al., 2018). Both models are maximum entropy graph models
(Garlaschelli et al., 2018), with sharp (all network realizations
satisfy the specified degree sequence) and soft (network realizations
satisfy the degree sequence on average) local constraints, respectively.
The (soft) configuration model is an unbiased random graph model widely
used by ecologists in the context of null hypothesis significance
testing of network structure (\emph{e.g.} Bascompte et al., 2003) and
can provide informative priors for Bayesian inference of network
structure (\emph{e.g.} J.-G. Young et al., 2021). It is noteworthy that
for this metaweb, the relevant information was extracted at the first
rank. Because the first rank corresponds to the leading singular value
of the system, the results of fig.~\ref{fig:degree} have a
straightforward interpretation: degree-based processes are the most
important in structuring the mammalian food web.

@@

One important aspect in which Europe and Canada differ (despite their
comparable bioclimatic conditions) is the degree of the legacy of human
impacts, which have been much longer in Europe. Nenzén et al. (2014)
showed that even at small scales (the Iberian peninsula), mammal food
webs retain the signal of both past climate change and human activity,
even when this human activity was orders of magnitude less important
than it is now. Similarly, Yeakel et al. (2014) showed that changes in
human occupation over several centuries can lead to food web collapse.
Megafauna in particular seems to be very sensitive to human arrival
(Pires et al., 2015). In short, there is well-substantiated support for
the idea that human footprint affects more than the risk of species
extinction (Marco et al., 2018), and can lead to changes in interaction
structure.

Cirtwill et al. (2019) showed that network inference techniques based on
Bayesian approaches would perform far better in the presence of an
interaction-level informative prior; the desirable properties of such a
prior would be that it is expressed as a probability, preferably
representing a Bernoulli event, the value of which would be
representative of relevant biological processes (probability of
predation in this case). We argue that the probability returned at the
very last step of our framework may serve as this informative prior;
indeed, the output of our analysis can be used in subsequent steps, also
possibly involving expert elicitation to validate some of the most
strongly recommended interactions. One important \emph{caveat} to keep
in mind when working with interaction inference is that interactions can
never really be true negatives (in the current state of our
methodological framework and data collection limitations); this renders
the task of validating a model through the usual application of binary
classification statistics very difficult (although see Strydom, Catchen,
et al., 2021 for a discussion of alternative suggestions). The other way
through which our framework can be improved is by substituting the
predictors that are used for transfer. For example, in the presence of
information on species traits that are known to be predictive of species
interactions, one might want to rely on functional rather than
phylogenetic distances -- in food webs, body size (and allometrically
related variables) has been established as such a variable (Brose et
al., 2006); the identification of relevant functional traits is
facilitated by recent methodological developments (Rosado et al., 2013).

Finally, it should be noted that the framework we have presented is
amenable to changes lending to applicability to a broad range of
potential scenarios. For example in this case study we have embedded the
original metaweb using t-SVD, because it lends itself to an RDPG
reconstruction, which is known to capture the consequences of
evolutionary processes (Dalla Riva \& Stouffer, 2016); this being said,
there are other ways to embed graphs (Arsov \& Mirceva, 2019; Cai et
al., 2017; Cao et al., 2019), which can be used as alternatives.
Regarding the transfer step it is possible to use distinct trees if
working with distinct clades (such as pollination networks) or an
alternative measure of similarity (transfer medium) such as information
on foraging (Beckerman et al., 2006), cell-level mechanisms (Boeckaerts
et al., 2021), or a combination of traits and phylogenetic structure
(Stock, 2021). Most importantly, although we focus on a trophic system,
it is an established fact that different (non-trophic) interactions do
themselves interact with and influence the outcome of trophic
interactions (see \emph{e.g.} Kawatsu et al., 2021; Kéfi et al., 2012).
Future development of metaweb inference techniques should cover the
prediction of multiple interaction types.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Albouy2019MarFis}{}%
Albouy, C., Archambault, P., Appeltans, W., Araújo, M. B., Beauchesne,
D., Cazelles, K., Cirtwill, A. R., Fortin, M.-J., Galiana, N., Leroux,
S. J., Pellissier, L., Poisot, T., Stouffer, D. B., Wood, S. A., \&
Gravel, D. (2019). The marine fish food web is globally connected.
\emph{Nature Ecology \& Evolution}, \emph{3}(8, 8), 1153--1161.
\url{https://doi.org/10.1038/s41559-019-0950-y}
